---
title: "Session 1"
author: "Raymond Hicks"
date: "May 15, 2018"
#output: html_document
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='c:/Users/arpie71/dropbox/columbia/workshop/data')
```

# Introduction to R
## Installing and loading packages

There are two ways to install packages. You can use the <b>install.packages</b> command. Or you can use the dropdown menu in RStudio.
The dropdown is Tools->Install Packages.
You only need to install packages once so it might make sense to use the dropdown.
In the first part of the lab, we will get R ready and install the relevant packages you will need.
R is annoying though and when you update it installs in a new directory. Sometimes the old packages need to be reinstalled using the new version. So updating can be a major task.

``` {r install and load packages}
#install.packages('quanteda')
# R ignores lines starting with a # 
# They are useful to add comments to files 
library(quanteda)

```

# Directories

Where are your files? 
You can see which directory you are in with the <b>getwd()</b> command.
<b>setwd()</b> changes the current directory. 
The <b>wd</b> parts of the commands are short for Working Directory.
The <b>list.files()</b> command will list the files in the working directory.

``` {r file locations}
getwd()
setwd('c:/Users/arpie71/dropbox/columbia/workshop/data')
list.files()

```

# Opening files
We will do most of the work in R using either comma-delimited (csv) or tab-delimited files (txt,tsv).
For both, we will use the <b>read.csv</b> command. 
The <i>header</i> suboption tells us whether the file has column names.
The <i>sep</i> suboption tells us what delimiter is used to separate columns of data.

``` {r use data}
setwd('c:/Users/arpie71/dropbox/columbia/workshop/data')
elec_data <- read.csv(file="2016_presidential_election_results.tsv", header=TRUE, sep="\t")
fs_data <- read.csv('2015_monthly_average_of_person_participating_in_food_stamp_program.tsv', header=TRUE, sep="\t")
income_data <- read.csv('2015_median_annual_hosehold_income.tsv', header=TRUE, sep="\t")
sa <- read.csv('state_abbreviations.tsv', header=TRUE, sep="\t")
census_data <- read.csv('population_by_state.tsv', header=TRUE, sep="\t")
income_data = read.csv('2015_median_annual_hosehold_income.tsv', header=TRUE, sep="\t")
trade76 <- read.csv(file="trade76.csv", header=TRUE, sep=",")

```

# Data
## investigating the data frame

``` {r investigate}

str(elec_data) # Structure of data
names(elec_data) # Column names
class(elec_data) # Format of data




```

# Looking at data
There are many different options to look at data. We will go over a few and their equivalencies.

## Top rows
There are a couple of ways to view the top rows. 
R has a <b>head</b> command similar to Python's <b>DF.head()</b> command.
We could also ask for the first N rows: elec_data[1:5,]

## Specific observations
There are different ways to view the contents of an individual row. 
The dataframe is like a matrix and we refer to observations within the matrix using square brackets ([]).
The first reference is to the row and the second to the column.
So we can ask R to return the row where State is equal to "New Jersey" by specifying the condition as the row option:
elec_data[elec_data$State=='New Jersey',]

``` {r datalook}
head(elec_data,5)
elec_data[1:5,]

# List  New Jersey data
elec_data$State=='New Jersey' # WRONG
elec_data[elec_data$State=='New Jersey',]
elec_data[31,]


```


# Specific columns within observations
The second part in the square brackets is the column information. If we leave it blank it returns all columns.
We can get a selected column by specifying its name.
elec_data[elec_data$State=='New Jersey','Sum']

For multiple columns, we need to use the combine feature in R.
This holds  a list or vector of names.
vlist<-c('Sum','Clinton','Trump')
Or we could specify it directly into the columns.

## Multiple rows

There are a couple of ways we could list multiple rows.
The combine feature is more like an "and" qualifier so it will not work with rows.
We could specifically ask for rows using the "|" symbol.
Or we could use the "in" option and the combine feature.
This tells R to list the row if State is in the list within combine.

``` {r mulitple obs}
# List specific columns for NJ 

elec_data[elec_data$State=='New Jersey','Sum']

# List multiple columns
vlist<-c('Sum','Clinton','Trump')
elec_data[elec_data$State=='New Jersey',vlist]

elec_data[elec_data$State=='New Jersey',c('Sum','Clinton','Trump')]

# List multiple rows

elec_data[(elec_data$State=='New Jersey'| elec_data$State=='New York'),c('Sum','Clinton','Trump')]
r <- c("New Jersey", "New York")
elec_data[elec_data$State %in% r ,c('Sum','Clinton','Trump')]
elec_data[elec_data$State %in% c("New Jersey","New York") ,c('Sum','Clinton','Trump')]


```




# Merging data
R's merge function is conveniently called <b>merge</b>. 
It's more useful than Stata's in some ways. You do not have to merge on variables with the same name. Instead, you could specify the merge variable separately for each dataset.
In the examples below if one of the datasets called the State variable StateName rather than State, we would pass that information to R:

df <- merge(elec_data, fs_data, by.x ='State', by.y= 'StateName')


``` {r merging}

df <- merge(elec_data, fs_data, by ='State')
dim(df) # equivalent to df.shape in Python
df <- merge(df, sa, by = 'State')
df <- merge(df, income_data, by='State')
df <- merge(df, census_data , by = 'State')
dim(df)

```

# Creating new columns
- Finally, we will briefly cover how to create new columns in R. 

-- We can create a new column based on data in existing columns.
--- df\$Trump_perc=df$Trump/(df\$Clinton+df\$Trump+df\$Johnson)
-- Or create new columns based on mathematical functions
--- elec_data\$numid<-seq(1:nrow(elec_data))
--- elec_data\$lnsum<-log(elec_data$Sum)



``` {r new columns}

df$Trump_perc=df$Trump/(df$Clinton+df$Trump+df$Johnson)
df$fs_perc=df$Participation/df$Population
df$vote_perc=(df$Clinton+df$Trump+df$Johnson)/df$Population
df$Trump_win=(df$Trump>df$Clinton)

# Sequence
df$numid<-seq(1:nrow(df))
#squared
df$numid2<-seq(1:nrow(df))^2

# Log of column
df$lnsum<-log(df$Sum)
```

# Save dataframe

- To save dataframe
-- save(df, file="elect.RData")

```{r save}
save(df, file="c:/Users/arpie71/Dropbox/Columbia/workshop/datafiles/elect.RData")
```
